---
title: "BIOL 606: Statistical Knowledge Assessment"
author: "Dr. Atkins"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```
## Instructions

This assessment helps me understand your statistical background. Please answer honestly—this is **not graded** and will help me tailor the course to your needs. Answer all questions to the best of your ability.

**Name:** ______________________________

**Email:** ______________________________





### Question 1

To the best of your ability, define the following terms:

&nbsp;&nbsp;a) Standard Error:
 
&nbsp;&nbsp;b) Variance:
 
&nbsp;&nbsp;c) Variability:

&nbsp;&nbsp;d) Covariance:
  
&nbsp;&nbsp;e) Central Tendency

\hfill\break

### Question 2

Provide an example of the following type of variables:

&nbsp;&nbsp;a) Discrete:
  
&nbsp;&nbsp;b) Continuous:
  
&nbsp;&nbsp;c) Binary:
  
&nbsp;&nbsp;d) Nominal:
  
&nbsp;&nbsp;e) Ordinal:

\hfill\break


### Question 3

How are parametric and non-parametric statistics different? 


\newpage

### Question 4

What does ANOVA stand for and how is it different than a T-test?

\hfill\break

### Question 5

Define the following terms:

&nbsp;&nbsp;a) Ordination:
  
&nbsp;&nbsp;b) Cluster Analysis:
  
&nbsp;&nbsp;c) Psuedoreplication:
  
&nbsp;&nbsp;d) Machine Learning:
  
&nbsp;&nbsp;e) Heteroskedasticity: 

\hfill\break

### Question 6

Interpret the plot below in as much detail as possible:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align = "center", fig.width = 4, fig.height = 3}
library(ggplot2)
library(dplyr)
data(InsectSprays)

# Calculate summary data
summary_data <- InsectSprays %>%
  group_by(spray) %>%
  summarize(
    mean_count = mean(count),
    se = sd(count) / sqrt(n()),
    .groups = "drop"
  )

# Run ANOVA and Tukey
anova_model <- aov(count ~ spray, data = InsectSprays)
tukey_results <- TukeyHSD(anova_model)

# Function to determine significance vs control (spray A)
get_sig_vs_control <- function(spray_name, tukey_results) {
  if (spray_name == "A") return("")  # Control
  
  comparison_name <- paste(spray_name, "A", sep = "-")
  p_value <- tukey_results$spray[comparison_name, "p adj"]
  
  if (p_value < 0.001) return("***")
  if (p_value < 0.01) return("**")
  if (p_value < 0.05) return("*")
  return("ns")
}

# Add significance symbols
summary_data$significance <- sapply(summary_data$spray, 
                                    get_sig_vs_control, 
                                    tukey_results = tukey_results)

# Create plot (remove x11())
ggplot(summary_data, aes(x = spray, y = mean_count, fill = spray)) +
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin = mean_count - se, 
                    ymax = mean_count + se),
                width = 0.3) +
  geom_text(aes(label = significance, y = mean_count + se + 1.5), 
            size = 6, fontface = "bold") +
  labs(x = "Insecticide Spray Type",
       y = "Mean Insect Count") +
  theme_bw() +
  theme(legend.position = "none")
```

\newpage

### Question 7

Interpret the plot below in as much detail as possible:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align = "center", fig.width = 6, fig.height = 6}
data(iris)
library(ggplot2)
library(ggbiplot)
iris.pca <- prcomp (~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width,
                    data=iris,
                    scale. = TRUE)

ggbiplot(iris.pca, obs.scale = 1, var.scale = 1,
           groups = iris$Species, point.size=2,
           varname.size = 5, 
           varname.color = "black",
           varname.adjust = 1.2,
           ellipse = TRUE, 
           circle = TRUE) +
labs(fill = "Species", color = "Species") +
theme_minimal(base_size = 14) +
theme(legend.direction = 'horizontal', legend.position = 'top')

```


\newpage

### Question 8

Interpret the plot below in as much detail as possible:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align = "center", fig.width = 6, fig.height = 6}
library(ggplot2)
library(ggdendro)

scaled_data <- scale(USArrests)
hc_ward <- hclust(dist(scaled_data), method="ward.D2")
ggdendrogram(hc_ward, rotate = TRUE, size = 2)
```


<br>
<br>
<br>
<br>

\newpage

### Question 9

What does statistical significance mean?

\hfill\break


### Question 10

Why did you take this class and what topics related to quantitative ecology are of most interest to you?

\hfill\break


### Question 11

List all of courses you have completed as either an undergraduate or graduate in the following areas (and please use descriptive titles rather than course numbers):

Ecology:
\hfill\break
\hfill\break
\hfill\break
Statistics:

\hfill\break

### Question 12

What is your coding experience? (any language)

<br>
<br>
<br>
<br>



